A possibilistic valid-time model

### 2011-12-01 ipmu2012/preliminaries.tex
A very small amend...

Both possibility measures and necessity measures are special cases of confidence measures.

@@

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set. In this case, the exact set is off course (partially) unknown, but the available knowledge is given in the form of a possibility distribution on the powerset $\Pow(R)$, which describes the possibility of each crisp subset of $R$ to be the value the variable takes. This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.

In this work, an ill-known set will be defined and represented via its start and end point, which will be ill-known values. The correspondences and transitions between these two representations are part of the authors current research.

@@

A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any boolean propositional variable $p$ takes just one value in $\mathbb{B}$. If the knowledge on which value this variable $p$ will take is given by a possibility distribution $\pi_p$, the variable can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demands most attention. This possibility and necessity is noted here as:
\begin{align}
Pos(p) & = \pi_p(T) \label{propholdsposs} \\
Nec(p) & = 1-\pi_p(F) \label{propholdsnecc}
\end{align}
Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}, instead of a \emph{fuzzy interval}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

A simple form of the membership function of a fuzzy interval is a trapezoidal function. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. A visualization is given in figure \ref{fig:trapezoidal}. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. The corresponding membership function definition for this $\mu_T$ is then given by:

\begin{align}
\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
 : & \quad x \rightarrow
\begin{cases}
1 & \mbox{ if } x \in [\beta,\gamma] \\
0 & \mbox{ if } x > \delta \vee x < \alpha \\
\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
\end{cases}
\end{align}

@@

\subsubsection{Interval Evaluation by Ill-known Constraints}
The problem of the interval evaluation is explained in \cite{Pon11}: for a crisp interval $I = \left[ a, b \right]$, the need exists to know if all points in this interval reside between the boundaries of the ill-known interval $\left[ X , Y \right]$, where $X,Y$ are ill-known values on the set of real numbers $\mathbb{R}$, and $\lambda$ is the evaluation function.

We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$. Therefore, we consider two ill-known constraints:

\begin{eqnarray}
C_1\stackrel{\triangle}{=}\left(\leq,X\right)\\
C_2\stackrel{\triangle}{=}\left(\geq,Y\right).
\end{eqnarray}

The following expressions compute the possibility and the necessity:

\begin{eqnarray}
\label{eq:interval-pos}
\Pos\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\sup_{a\geq w}\pi_{X}(w),\sup_{b\leq w}\pi_{Y}(w)\bigg)\\
\label{eq:interval-nec}
\Nec\left(\lambda([a,b])\right)=\\
\nonumber
\min\bigg(\inf_{a<w}1-\pi_{X}(w),\inf_{b>w}1-\pi_{Y}(w)\bigg).
\end{eqnarray}

>>>>

%Both possibility measures and necessity measures are special cases of confidence measures.

@@

When a possibilistic variable is defined on the powerset $\Pow(R)$ of some universe $R$, the unique value the variable takes will be a crisp set and its possibility distribution on the powerset $\Pow(R)$ will describe the possibility of each crisp subset of $R$ to be the value the variable takes. This exact value (a crisp set) the variable takes, is now called an \emph{ill-known set}\cite{Dubois88b}.

@@

A specific application of possibilistic variables is obtained when the universe under consideration is the set of boolean values $\mathbb{B}$ = $\{T,F\}$. Indeed, any boolean proposition $p$ takes just one value in $\mathbb{B}$. If the knowledge on which value this proposition $p$ will take is given by a possibility distribution $\pi_p$, the proposition can be seen as a possibilistic variable. As the interest lies with the case where the proposition holds, the possibility and necessity that $p$ = $T$ (the proposition holds) demand most attention. This possibility and necessity is noted here as:
\begin{align}
\text{Possibility that $p$ = $T$ (p holds):} \hspace{50pt} & Pos(p) = \pi_p(T) \label{propholdsposs} \\
\text{Necessity that $p$ = $T$ (p holds):} \hspace{50pt} & Nec(p) = 1-\pi_p(F) \label{propholdsnecc}
\end{align}
%Here, notation \ref{propholdsposs} denotes the possibility that $p$ = $T$ and the proposition holds, notation \ref{propholdsnecc} denotes the necessity that $p$ = $T$ and the proposition holds.

This work will deal with ill-known intervals. These are ill-known sets, defined and represented via a start and end point, which will be ill-known values. The elements of the set are the values between the start and end point. An closed ill-known interval with start point $X$ and end point $Y$ is noted here $\left[X, Y\right]$. The correspondences and transitions between the representations of ill-known sets and between the representation of an ill-known set and an ill-known interval are part of the authors current research.

\subsection{\label{subsec:fuzzy-numbers}Fuzzy Numbers and Fuzzy Intervals}
Among others, Dubois and Prade~\cite{Dubois1983} use the following definition of a \emph{fuzzy interval}.
\begin{definition}
A fuzzy interval is a fuzzy set $M$, defined by a membership function $\mu_{M}$, on the set of real numbers $\mathbb{R}$ such that:
\begin{eqnarray}
\mu_{M} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mathbb{R} \rightarrow \left[0,1\right] \nonumber \\ 
\forall (u,v)\in\mathbb{R}^2: \forall w \in [u,v]:&\mu_M(w) \geq\min(\mu_M(u),\mu_M(v))  \\
\exists m \in \mathbb{R} : & \!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\!\! \mu_M(m)=1 
\end{eqnarray}
\end{definition}
If this modal value $m$ is unique, then $M$ is referred to as a \emph{fuzzy number}. In other words, if the core of a fuzzy interval is a singleton, it is referred to as a fuzzy number.

A simple form of the membership function of a fuzzy interval is a trapezoidal function. It can be shown that such a membership function $\mu_T$ for a fuzzy interval $T$ is convex and normalized. A visualization is given in figure \ref{fig:trapezoidal}. Four reel values, denoted $\alpha$, $\beta$, $\gamma$ and $\delta$ and chosen as in figure \ref{fig:trapezoidal}, suffice to represent a trapezoidal membership function of a fuzzy interval. In this work, a fuzzy interval defined as such will be noted as $\left[\alpha, \beta, \gamma, \delta\right]$. %The corresponding membership function definition for this $\mu_T$ is then given by:

%\begin{align}
%\mu_T : & \quad \mathbb{R} \rightarrow \left[0,1\right] \\
% : & \quad x \rightarrow
%\begin{cases}
%1 & \mbox{ if } x \in [\beta,\gamma] \\
%0 & \mbox{ if } x > \delta \vee x < \alpha \\
%\frac{x-\alpha}{\beta - \alpha} & \mbox{ if } x \in [\alpha,\beta[ \\
%\frac{\delta -x}{\delta - \gamma} & \mbox{ if } x \in ]\gamma,\delta] \\
%\end{cases}
%\end{align}

@@

\subsubsection{Interval Evaluation by Ill-known Constraints}
The problem of interval evaluation is more generally explained in \cite{Pon11}: the need exists to know if all points in a crisp interval $I$ reside between the boundaries of an ill-known interval $\left[ X , Y \right]$. In \cite{Pon11}, the notion of an \emph{ill-known constraint} is introduced:

\begin{definition}
Given a universe $U$, an \emph{ill-known constraint} $C$ on a set $A \subseteq U$ is specified by means of a binary relation $R \subseteq U^{2}$ and a fixed, ill-known value denoted by its possibilistic variable $V$ over $U$, i.e.:
\begin{align}
C \triangleq (R,V)
\end{align}
Set $A$ now satisfies the constraint if and only if:
\begin{align}
\forall a \in A : (a,V) \in R
\end{align}
\end{definition}

The satisfaction of a constraint $C \triangleq (R,V)$ by a set $A$ is now still a Boolean matter, but due to the uncertainty about the ill-known value $V$, it can be uncertain whether $C$ is satisfied by $A$ or not\cite{Pon11}. In fact, this satisfaction now behaves as a proposition. Based on the possibility distribution $\pi_{V}$ of $V$, the possibility and necessity that $A$ satisfies $C$ can be found. This proposition can thus be seen as a possibilistic variable on $\mathbb{B}$. The required possibility and necessity are:

\begin{align}
\Pos(A\text{ satisfies }C) & = \min_{a \in A}\left(\sup_{(a,w) \in R}\pi_{V}(w))\right) \\
\Nec(A\text{ satisfies }C) & = \min_{a \in A}\left(\inf_{(a,w) \notin R} 1-\pi_{V}(w)\right)
\end{align}

Now, to check if crisp interval $I = \left[j, k\right]$ is included in $\left[X, Y\right]$, 2 ill-known constraints are constructed:

%We assume that $X$ specifies the lower bound and $Y$ the upper bound for a given interval, we want to known whether all points in the interval are larger than or equal to $X$ and smaller than or equal to $Y$. Therefore, we consider two ill-known constraints:

\begin{eqnarray}
C_1\triangleq\left(\geq,X\right)\\
C_2\triangleq\left(\leq,Y\right).
\end{eqnarray}

The possibility and necessity of the set satisfying both constraints is then:

\begin{align}
\label{eq:interval-pos}
\Pos(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\sup_{a \geq w}\pi_{X}(w),\sup_{a \leq v}\pi_{Y}(v)\right)\\
\label{eq:interval-nec}
\Nec(A\text{ satisfies }C_1\ AND\ C_2) & = \min_{a \in A}\left(\inf_{a < w} 1-\pi_{X}(w),\inf_{a > v} 1-\pi_{Y}(v)\right).
\end{align}
